High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-10-07 inst/manuscript/figures.Rmd
Revisions for Frontiers in AI round 1 reviews

### Figure 1: System architecture
Created in keynote

@@

### Figure S3: Map reference sample
```{r, eval = FALSE}
ref_sample <- readr::read_csv(
  system.file("extdata", "map_reference_sample.csv", package = "activemapper")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

@@

### Figure S4: AOI Zones

@@

# read in field segments
flds_1 <- read_sf(segment_files$file[1])
flds_5 <- read_sf(segment_files$file[5])
flds_11 <- read_sf(segment_files$file[11])
flds_15 <- read_sf(segment_files$file[15])

@@

# process inset maps
# img <- read_stars(impath)
# mxval <- st_apply(img, 3, max) %>% pull(max)
# imgngb <- st_rgb(img[, , , 4:2], 3, maxColorValue = mxval[4:2])
vpal <- viridis::viridis(20)
inset_cols <- c("red", "orange", "pink", "brown")
fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
isets <- lapply(1:4, function(x) {  # x <- 1
  imgr <- brick(impath[x])[[2:4]]
  probr <- stars::read_stars(prob_tiles$impath[x])
  xylims <- st_bbox(imgr)
  iset <- ggplot() + 
    ggRGB(imgr, maxpixels = 50000, stretch = "lin", ggLayer = TRUE) +
    geom_sf(data = fld_tiles[[x]], fill = "transparent", color = "yellow", 
            size = 0.2) + 
    geom_sf(data = tile_ss[x, ], color = inset_cols[x], fill = "transparent") + 
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void()
  iset2 <- ggplot() + 
    geom_stars(data = probr) +
    scale_fill_gradientn(name = "", colours = vpal, limits = c(0, 100)) +
    geom_sf(data = tile_ss[x, ],  color = inset_cols[x], fill = "transparent") +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void() + theme(legend.position = "none")
  cowplot::plot_grid(iset2, iset, ncol = 1, align = "hv")
})
# isets[[2]]

@@

# cropland fractions
p <- ggplot(africa) + 
  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
  geom_stars(data = cropfrac) +
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
  geom_sf(data = tile_ss, fill = inset_cols, col = inset_cols, size = 0.1) +
  # geom_sf(data = tile_ss, fill = "red", col = "red") +
  scale_fill_viridis_c(name = "%", na.value = "transparent") +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_void() + 
  theme(legend.position = c(0.64, 0.08), legend.direction = "horizontal", 
        legend.key.height = unit(0.3, "cm"), 
        legend.text = element_text(colour = "black"),
        legend.title = element_text(size = 10, colour = "black"))
# with insets added
wh <- c(0.4, 0.4)
lx <- -0.1
rx <- 0.7
inset_map <- ggdraw() +
  draw_plot(p) +
  draw_plot(isets[[1]], x = lx, y = 0.6, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[2]], x = rx, y = 0.51, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[3]], x = lx, y = 0.17, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[4]], x = rx, y = 0.07, width = wh[1], height = wh[2])
# inset_map

@@

### Figure S5: PlanetScope image density
The density of imagery available for download from the image catalog.

@@

### Figure S6: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

@@

### Figure S7: Assignment count
```{r assignmentcount, eval=FALSE}
data("label_summary")

@@

### Figure S9: Label quality
```{r labelquality, eval=FALSE}
# Plot data from Figure S7

@@

### Figures S10 and S11: Consensus label (Bayesian) risk

@@

### Figure S12: Training data error impact on probability images
```{r, eval = FALSE}
probpal <- viridis::viridis
cprob <- raster(here::here("external/prob/image_c294_r532.tif"))
hprob <- raster(here::here("external/prob/image_c294_r532_high.tif"))
lprob <- raster(here::here("external/prob/image_c294_r532_low.tif"))

@@

### Figures S13 and S14: Segmentation statistics (grid scale)

@@

### Figure S15: Average fields sizes

>>>>

### Figure 1: System architecture
Created in Keynote

@@

### Figure S3: Segmentation workflow 
Created in Keynote.

### Figure S4: Map reference sample
```{r, eval = FALSE}
ref_sample <- readr::read_csv(
  system.file("extdata", "map_reference_sample.csv", package = "activemapper")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

@@

### Figure S5: AOI Zones

@@

# read in field segments
flds_1 <- read_sf(segment_files$file[1]) %>% st_make_valid()
flds_5 <- read_sf(segment_files$file[5]) %>% st_make_valid()
flds_11 <- read_sf(segment_files$file[11]) %>% st_make_valid()
flds_15 <- read_sf(segment_files$file[15]) %>% st_make_valid()

@@

# process inset maps
# img <- read_stars(impath)
# mxval <- st_apply(img, 3, max) %>% pull(max)
# imgngb <- st_rgb(img[, , , 4:2], 3, maxColorValue = mxval[4:2])
vpal <- viridis::viridis(20)
# inset_cols <- c("red", "orange", "pink", "brown")
inset_cols <- rep("red", 4)
fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
isets <- lapply(1:4, function(x) {  # x <- 1
  imgr <- brick(impath[x])[[2:4]]
  probr <- stars::read_stars(prob_tiles$impath[x])
  xylims <- st_bbox(imgr)
  iset <- ggplot() + 
    ggRGB(imgr, maxpixels = 50000, stretch = "lin", ggLayer = TRUE) +
    geom_sf(data = fld_tiles[[x]], fill = "transparent", color = "yellow", 
            size = 0.2) + 
    geom_sf(data = tile_ss[x, ], color = inset_cols[x], fill = "transparent") + 
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void()
  iset2 <- ggplot() + 
    geom_stars(data = probr) +
    scale_fill_gradientn(name = "", colours = vpal, limits = c(0, 100)) +
    geom_sf(data = tile_ss[x, ],  color = inset_cols[x], fill = "transparent") +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void() + theme(legend.position = "none")
  cowplot::plot_grid(iset2, iset, ncol = 1, align = "hv")
})
# isets[[2]]

@@

# create lines to point to insets
tile_ss_xy <- st_coordinates(st_centroid(tile_ss))
lns <- st_sfc(
  list(
    st_linestring(
      rbind(tile_ss_xy[1, ], c(maprange[1] * 0.9, maprange[4] * 0.885))
    ), 
    st_linestring(
      rbind(tile_ss_xy[2, ], c(maprange[2] * 0.9, maprange[4] * 0.8517))
    ),
    st_linestring(
      rbind(tile_ss_xy[3, ], c(maprange[1] * 0.9, maprange[4] * 0.619))
    ),
    st_linestring(
      rbind(
        tile_ss_xy[4, ], c(maprange[2] * 0.9, maprange[4] * 0.5681))
    )
  ), crs = 4326
)

# cropland fractions
p <- ggplot(africa) + 
  geom_sf(fill = "gray60", color = "grey", size = 0.25) +  
  geom_stars(data = cropfrac) +
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
  geom_sf(data = tile_ss, fill = inset_cols, col = inset_cols, size = 0.1) +
  geom_sf(data = lns, col = inset_cols) + 
  geom_sf(data = tile_ss, fill = "red", col = "red") +
  scale_fill_viridis_c(name = "%", na.value = "transparent") +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  ggsn::north(symbol = 16, scale = 1.2, x.min = 0.5, x.max = 1, 
              y.min = 10.5, y.max = 11) +
  ggsn::scalebar(ghana, dist = 100, dist_unit = "km", location = "bottomleft", 
                 transform = TRUE, model = "WGS84", st.size = 3, st.dist = 0.01, 
                 anchor = c(x = -0.7, y = 4.8), border.size = 0.1) + 
  theme_void() + 
  theme(legend.position = c(0.675, 0.1), legend.direction = "horizontal",
        legend.key.height = unit(0.3, "cm"),
        legend.key.width = unit(0.67, "cm"),
        legend.text = element_text(colour = "black"),
        legend.text.align = 0.5,
        legend.title.align = 1,
        legend.spacing.y = unit(0.05, "cm"),
        legend.title = element_text(
          vjust = 1, size = 11, colour = "black"
        )
  )

# with insets added
wh <- c(0.4, 0.4)
lx <- -0.1
rx <- 0.7
inset_map <- ggdraw() +
  draw_plot(p) +
  draw_plot(isets[[1]], x = lx, y = 0.6, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[2]], x = rx, y = 0.51, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[3]], x = lx, y = 0.17, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[4]], x = rx, y = 0.07, width = wh[1], height = wh[2])
# inset_map

@@

### Figure S6: PlanetScope image density
The density of imagery available for download from the image catalog.

@@

### Figure S7: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

@@

### Figure S8: Assignment count
```{r assignmentcount, eval=FALSE}
data("label_summary")

@@

### Figure S10: Label quality
```{r labelquality, eval=FALSE}
# Plot data from Figure S7

@@

### Figures S11 and S12: Consensus label (Bayesian) risk

@@

### Figure S13: Training data error impact on probability images
```{r, eval = FALSE}
probpal <- viridis::viridis
cprob <- raster(here::here("external/prob/image_c294_r532.tif"))
hprob <- raster(here::here("external/prob/image_c294_r532_high.tif"))
lprob <- raster(here::here("external/prob/image_c294_r532_low.tif"))

@@

### Figures S14 and S15: Segmentation statistics (grid scale)

@@

### Figure S16: Average field sizes

### 2021-10-07 inst/manuscript/si.Rmd
Revisions for Frontiers in AI round 1 reviews

>>>>

### Segmentation
An overview of the two inputs to and outputs from the segmentation algorithm are shown in Figure \ref{fig:segfig}.

```{r segfig, echo = FALSE, fig.align='center', out.width="100%", message=FALSE, warning=FALSE, fig.cap="An overview of the inputs and key outputs from the five-step segmentation algorithm."}
knitr::include_graphics('figures/si_segmentation_workflow.png')
```

### 2021-10-07 inst/manuscript/si.tex
Revisions for Frontiers in AI round 1 reviews

@@

The average Bayesian Risk of each training and validation site is shown
in Figure \ref{fig:labelrisk}, as well as the distribution of risk
values per AOI and the initial training clusters (Figure
\ref{fig:labelriskhist}). The three initial clusters include the second
mapping of Cluster 2 (Figure \ref{fig:trainval}A).

@@

The average Bayesian risk was 0.122 for training labels and 0.127 for
validation labels. Risk was highest in the northern AOIs (AOIs 1-6;
Figures S6-7), falling between 0.157 for training and 0.173 for
validation labels (Figures S6-7), and lowest in the southwestern AOIs
(AOIs 10, 11, 13, 14, 16; training risk = 0.079; validation risk =
0.065). Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15)
was slightly lower (training = 0.127; validation = 0.136) than in the
north. Labeller experience also appeared to reduce risk, which we
observed during a relabelling of the 500 initial random site in this
cluster; the mean risk of the updated labels was 0.055, compared to
0.172 for original labels.

@@

\begin{itemize}
\tightlist
\item
  \href{https://github.com/agroimpacts/activemapper}{\texttt{activemapper}}:
  Contains the source code and derived data for this paper. The
  repository\footnote{github.com/agroimpacts/activemapper} will be
  opened following publication in a peer-reviewed journal.
\item
  \href{https://github.com/agroimpacts/croplandmaps_V2}{\texttt{croplandmaps}}:
  Source code for web map displaying image composites and cropland
  probability maps. The
  repository\footnote{github.com/agroimpacts/croplandmaps\_V2} is open,
  and the field boundary datasets are available for download.
\item
  \href{https:github.com/agroimpacts/imager}{\texttt{imager}}: Source
  code for the image processing software. The
  repository\footnote{github.com/agroimpacts/imager} is currently
  private, but will be opened after completing a security audit.
\item
  \href{https:github.com/agroimpacts/labeller}{\texttt{labeller}}:
  Source code for the labelling platform. The
  repository\footnote{github.com/agroimpacts/labeller} is currently
  private, but will be opened after completing a security audit.
\item
  \href{https:github.com/agroimpacts/learner}{\texttt{learner}}: Source
  code for the active learning component. The
  repository\footnote{github.com/agroimpacts/learner} is open.
\item
  \href{https:github.com/agroimpacts/segmenter}{\texttt{segmenter}}:
  Source code for the segmentation algorithm. The repository
  \footnote{github.com/agroimpacts/segmenter} is currently private, but
  will be opened after completing a security audit.
\end{itemize}

@@

\leavevmode\hypertarget{ref-hijmansRasterGeographicData2020}{}%
Hijmans, R. J. 2020. Raster: {Geographic} data analysis and modeling.
Manual.

\leavevmode\hypertarget{ref-pebesmaSimpleFeaturesStandardized2018}{}%
Pebesma, E. 2018. Simple features for r: {Standardized} support for
spatial vector data. The R Journal 10:439--446.

>>>>

\hypertarget{segmentation}{%
\subsubsection{Segmentation}\label{segmentation}}

An overview of the two inputs to and outputs from the segmentation
algorithm are shown in Figure \ref{fig:segfig}.

{\centering \includegraphics[width=1\linewidth,]{figures/si_segmentation_workflow}

\caption{An overview of the inputs and key outputs from the five-step segmentation algorithm.}\label{fig:segfig}
\end{figure}

@@

The average Bayesian Risk of each training and validation site is shown
in Figure \ref{fig:labelrisk}, and the distribution of risk values per
AOI and the initial training clusters in Figure \ref{fig:labelriskhist}.
The three initial clusters include the second mapping of Cluster 2
(Figure \ref{fig:trainval}A).

@@

The average Bayesian risk was 0.122 for training labels and 0.127 for
validation labels. Risk was highest in the northern AOIs (AOIs 1-6;
Figures \ref{fig:labelrisk}-\ref{fig:labelriskhist}), falling between
0.157 for training and 0.173 for validation labels, and lowest in the
southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = 0.079;
validation risk = 0.065). Label risk in the central-southeastern AOIs
(AOIs 7-9, 12, 15) was slightly lower (training = 0.127; validation =
0.136) than in the north. Labeller experience also appeared to reduce
risk, which we observed during a relabelling of the 500 initial random
site in this cluster; the mean risk of the updated labels was 0.055,
compared to 0.172 for original labels.

@@

\begin{itemize}
\tightlist
\item
  \href{https://github.com/agroimpacts/activemapper}{\texttt{activemapper}}:
  Contains the source code and derived data for this paper. The
  repository\footnote{github.com/agroimpacts/activemapper} will be
  opened following publication in a peer-reviewed journal.
\item
  \href{https://github.com/agroimpacts/croplandmaps_V2}{\texttt{croplandmaps}}:
  Source code for web map displaying image composites and cropland
  probability maps. The
  repository\footnote{github.com/agroimpacts/croplandmaps\_V2} is open,
  and the field boundary datasets are available for download.
\item
  \href{https:github.com/agroimpacts/imager}{\texttt{imager}}: Source
  code for the image processing software. The
  repository\footnote{github.com/agroimpacts/imager} is currently
  private, but will be opened after completing a security audit.
\item
  \href{https:github.com/agroimpacts/labeller}{\texttt{labeller}}:
  Source code for the labelling platform. The
  repository\footnote{github.com/agroimpacts/labeller} is currently
  private, but will be opened after completing a security audit.
\item
  \href{https:github.com/agroimpacts/learner}{\texttt{learner}}: Source
  code for the active learning component. The
  repository\footnote{github.com/agroimpacts/learner} is open.
\item
  \href{https:github.com/agroimpacts/segmenter}{\texttt{segmenter}}:
  Source code for the segmentation algorithm. The repository
  \footnote{github.com/agroimpacts/segmenter} is open.
\end{itemize}

@@

\leavevmode\hypertarget{ref-hijmansRasterGeographicData2020}{}%
Hijmans, R. J. 2020. Raster: Geographic data analysis and modeling.
Manual.

\leavevmode\hypertarget{ref-pebesmaSimpleFeaturesStandardized2018}{}%
Pebesma, E. 2018. Simple features for r: Standardized support for
spatial vector data. The R Journal 10:439--446.
